---                                                                    
title: "Getting into flow cyto analysis in R, basics"
author: "Darach Miller"                                                   
date: "`r Sys.Date()`"                                                    
output:                                                                   
  html_document:                                                          
    self_contained: true                                                  
---                           

I believe you should compartamentalize your reading in from files
and your analysis. You should use the below script is just to 
yank it in, and spit it out as a CSV.

Below is just the directory where my FCS files are. 

```{r}
directories <- c(
    exp154sytoxgreen="/home/zed/lab/data/mine/1601/exp154sytoxgreen/"
  )
directories
dir(directories[1])
```

See how that works? Change it to where your stuff is. Name it
according to the experiment name.


Here we immediately use `flowCore` just to extract the measurements 
from the machine. Note the lack of any confounding transformation.
This is 2016, and I'm pretty sure these are linear digital detectors.

```{r}
require(flowCore)
datz <- list()
for (expr in names(directories)) {
  dirz <- directories[expr]
  for (fz in dir(dirz)[grep("\\.fcs$",dir(dirz))]) { 
    nameToSaveAs <- paste0(expr, warning=F,"_",sub(".fcs","",fz))
    print(paste0("doing ",dirz,"",fz,", saving it as ",nameToSaveAs))
    datz[[nameToSaveAs]] <- data.frame(
        exprs(read.FCS(paste0(dirz,fz),transformation=FALSE))
      )
  }
}
```

The below takes everything in `datz` and `rbind()`'s it together
to make one large data frame. This might take a while.

```{r}
if (exists("lrgdf")) {rm(lrgdf)}
for (fz in names(datz)) {
  relevantBit <- gsub(".+?\\s(.+?)$","\\1",fz)
  tmpdf <- data.frame(datz[[as.character(fz)]],
    colsplit(relevantBit,"_",names=c("sample","dye","pretreatment")))
  if (!exists("lrgdf")) { lrgdf <- tmpdf } 
  else { lrgdf <- rbind(lrgdf,tmpdf) }
}
dim(lrgdf)
head(lrgdf)
```

And you can save that as a csv if you'd like:

```{r}
write.csv(lrgdf,file=paste0(names(directories)[1],"160510.csv"),
  quote=F,row.names=F)
```

Since I've got your attention, I'm going to show you how to gate 
like a neadertal. It works.
Note that I'm plotting a random subset of rows in order to make this 
plot more quickly.

```{r}
aSubsetOflrgdf <- lrgdf[sample(1:nrow(lrgdf),1e4),]
g <- ggplot(aSubsetOflrgdf)+
  aes(x=log(FSC.A),y=log(FSC.H))+
  geom_point(size=.5,alpha=.05)
g
g+geom_vline(xintercept=10.5)+geom_abline(slope=0.9,intercept=1.6)
```

Looks like singlets.


```{r}
subdf<-subset(lrgdf,log(FSC.A)>10.5&(log(FSC.H)>log(FSC.A)*0.9+1.6))
aSubsetOfsubdf <- subdf[sample(1:nrow(subdf),1e4),]
```

```{r}
g <- ggplot(aSubsetOfsubdf)+
  aes(x=log(FSC.A),y=log(FL1.A))+
  geom_point(size=.5,alpha=.05)+
  facet_wrap(~sample)
g
g+geom_hline(yintercept=13)
```

Same thing.

```{r}
subdf <- subset(subdf,log(FL1.A)>13)
aSubsetOfsubdf <- subdf[sample(1:nrow(subdf),1e4),]
```

```{r}
g <- ggplot(aSubsetOfsubdf)+
  aes(x=log(FSC.A),y=log(FL1.A))+
  geom_point(size=.5,alpha=.05)+
  facet_wrap(~sample)
g+coord_cartesian(xlim=c(10.5,14.5),ylim=c(13,15))
```

Close enoguht, although you can see some contamination. Let's look
at that in a different way, this time for all datar.

```{r}
subdf$Machine <- ifelse(grepl("g",subdf$sample),
  "chemostat1","chemostat2")
subdf$Limitation <- ifelse(grepl(".[123]",subdf$sample),
  "Limitation1","Limitation2")
subdf$Rate <- rep(c("slow","medium","fast"),2)[
  as.numeric(sub("^.","",subdf$sample))]
g <- ggplot(subdf)+aes(x=log10(FL1.A),col=Rate)+
  stat_bin(aes(y = ..density..),binwidth=0.005,geom="line",
    position="identity",size=.5)+
  stat_bin(aes(y = ..density..),binwidth=0.005,geom="point",
    position="identity",size=.5)+
  coord_cartesian(xlim=c(5.9,6.6))+
  theme(legend.position="bottom")
g+facet_grid(Machine~Limitation)
g%+%subset(subdf,Limitation=="Limitation1"&Rate=="fast"&Machine=="chemostat2")+
  guides(col=F)
```

That's text book.

One of these days, I'll write up how to use `mle2` to fit a mixed
model to estimate cell cycle contribution.

Anyone want to contribute and do this one?

